In this thesis, we have presented our implementation of convolutional neural networks that use the \textit{Depresjon} dataset, which is a minute by minute log of motor activity for each participant. We created three models, and the first can classify with promising performance (F1-score of 0.70) whether a participant belongs to the condition group (bipolar and unipolar depressed patients) or the control group (healthy participants). 

Another convolutional model, with the same data as input, detects one of three different levels of depression based on MADRS scores. We labeled participants in the control group as non-depressed and divided participants in the condition group into mildly depressed (MADRS between 7 and 19) and moderately depressed (MADRS between 20 and 34). Then we trained the model to detect this label for participants. We achieved an overall F1-score of 0.3 for this goal, which has a significant room for improvement.  

Our third and last model is a prediction model that aims to predict the MADRS score of participants, again using the same motor activity dataset as input. We did not leave one participant out to test on as we did in the other goals, as we did not have the computing hardware to train the model 55 times. Instead, we trained one model for 2700 epochs and achieved a mean squared error of approximately 4.0. 

We found that our models performed almost flawlessly (F1-scores above 99\% for classification and mean squared error of approximately 4) when training on data that includes the tested participants. The difference between these results and the experiments where we left one participant out of training is interesting, and it may indicate that convolutional networks are more useful for personal datasets. 

We presented both a medical and technical background in chapter 2, in addition to mentioning related work to our thesis. We described both bipolar disorder and Montgomery-Åsberg Depression Rating Scale as part of the medical background. It was essential to gain knowledge about these topics because as students in Computer Science, our knowledge was limited. The technical background contains an introduction to machine learning, which is both theoretical and practical. In both the examples in this part and our model implementations, we used Keras, a machine learning framework in Python, because of its simplicity. 

In the next chapter, we described the dataset, our primary goals and how we structured our data so that a convolutional neural network would be able to learn from it. The input data consisted of time-sliced segments mapped to the corresponding element in the output data. The output data contained one of three target values (from the demographics part of the dataset), which depended on the goal. We also touched upon different performance metrics that we used later in classification experiments.

In chapter 4, we introduced a small regression test to see if we could learn something from any column in the demographics dataset. As we expected, only the MADRS score had a relation to whether a participant was in the control or condition group. Then we proceeded to implement our convolutional neural networks. We built three different models for our three primary goals (the first two models were very similar, as only the number of output possibilities changed - depression level instead of control/condition). The last model was different in several layers, as it was built to predict MADRS values. 

We presented the training results in chapter 5. For each goal (including linear regression), we described the hyper-parameters we used when training the models. For the convolutional neural networks, we started out finding the optimal segment length for the input data before we trained the models. Then we calculated performance scores for the trained models and did cross-validation to ensure the consistency of the models. For the classifiers, we also did a final experiment where we left participants out one by one before training the models, then tested the models' capabilities to detect that particular participant.

We discussed our work in chapter 6 and came up with several issues that made our models perform the way they did. The amount of participants is limited, which is the main issue that we think reduces the detection capabilities of the models when it comes to testing on completely untouched participants. Other possible improvements are experimenting more with the segment lengths (input data) and other hyper-parameters. 

We compared our work to the research of Garcia-Ceja, E. et al. performed on the same dataset and found that the difference between their decision tree and neural network and our convolutional network was not that significant as we thought it would be. We discussed real-world use cases for convolutional neural networks in mental health. The excellent performance from when we included data from all participants in training gave us the idea that this kind of machine learning would perform better in detection within personal datasets. We also discussed a few ethical concerns that arise in the field of Mental Health Monitoring Systems. <More>

\section{Future Research}
